---
title: "Report for project 1 on nearest neighbors"
author: "Allen, Joe, Madi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Report for project 1 on nearest neighbors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r data}
data.name.vec <- c(
  "spam",
  "SAheart",
  "zip.train",
  "prostate",
  "ozone")
data.list <- list()
library(ElemStatLearn)
for(data.name in data.name.vec){
  data(list=data.name, package="ElemStatLearn")
  data.list[[data.name]]<-get(data.name)
}
is.binary <- ElemStatLearn::zip.train[,1] %in% c(0,1)
data.list <- list(
  spam=list(
    label.vec=ifelse(ElemStatLearn::spam$spam=="spam",1,0),
    feature.mat=as.matrix(ElemStatLearn::spam[, -ncol(ElemStatLearn::spam)])),
  zip.train=list(
    label.vec=ElemStatLearn::zip.train[is.binary,1],
    feature.mat=ElemStatLearn::zip.train[is.binary,-1]
  ))
n.folds <- 5
results.list <- list()
for(data.name in names(data.list)){
  one.data.set <- data.list[[data.name]]
  set.seed(1)
  fold.vec <- sample(rep(1:n.folds,l=nrow(one.data.set$feature.mat)))
  for(test.fold in 1:n.folds){
    is.train <- fold.vec != test.fold
    fit <- NNLearnCV(
    one.data.set$feature.mat[is.train,],
    one.data.set$label.vec[is.train])
    pred.vec<-fit$predict(one.data.set$feature.mat[!is.train,])
    ## compute test error of pred.vec with respect to one.data.set$label.vec[!is.train]
    results.list[[paste(data.name, test.fold)]]
  }
}

```
## Data set 1: spam

### Matrix of loss values

print out and/or plot the matrix.

comment on difference between NN and baseline.

### Train/validation loss plot

plot the two loss functions.

What is the optimal number of neighbors?

## Data set 2: SAheart

### Matrix of loss values

print out and/or plot the matrix.

comment on difference between NN and baseline.

### Train/validation loss plot

plot the two loss functions.

What is the optimal number of neighbors?


Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
