---
title: "Report for project 1 on nearest neighbors"
author: "Allen, Joe, Madi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Report for project 1 on nearest neighbors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r data}
data(zip.train, package="ElemStatLearn")
  i01 <- which(zip.train[,1] %in% c(0,1))
  train.i <- i01[1:5]
  test.i <- i01[6]
  x <- zip.train[train.i, -1]
  y <- zip.train[train.i, 1]
  testx.vec <- zip.train[test.i, -1]
  max.neighbors <- 4
  results.vec <- knn(x, y, testx.vec, max.neighbors)

```
## Data set 1: spam

### Matrix of loss values
```{r}
results <- results.list[["spam"]]
```

#### Plot
```{r}
plot(results)
```

comment on difference between NN and baseline.

### Train/validation loss plot

```{r}
results <- results.list[["zip.train"]]
```

What is the optimal number of neighbors?

## Data set 2: SAheart

### Matrix of loss values

```{r}
results <- results.list[["SAheart"]]

```

#### Plot
```{r}
plot(results)
```

comment on difference between NN and baseline.

### Train/validation loss plot

```{r}
results <- results.list[["zip.train"]]
```

What is the optimal number of neighbors?

## Data set 3: prostate

### Matrix of loss values

```{r}
results <- results.list[["prostate"]]
```

#### Plot
```{r}
plot(results)
```

comment on difference between NN and baseline.

### Train/validation loss plot

```{r}
results <- results.list[["zip.train"]]
```

What is the optimal number of neighbors?

## Data set 4: ozone

### Matrix of loss values

```{r}
results <- results.list[["ozone"]]
plot(results)
```

#### Plot
```{r}
plot(results)
```

comment on difference between NN and baseline.

### Train/validation loss plot

```{r}
results <- results.list[["zip.train"]]
```

What is the optimal number of neighbors?
